Cyclone: High Availability for Persistent Key Value Stores

### 2017-12-21 paper/paper.tex
Switch to usenix format.
Update.

\documentclass[twocolumn]{article}

@@

\title{Cyclone: High Availability for Persistent Key Value Stores}
\author{Amitabha Roy \\ amitabha.roy@gmail.com 
	     \and Subramanya R. Dulloor \\ dulloor@gmail.com}

@@

\begin{abstract}
\vspace{0.05in}
Persistent key value stores are an important component of many distributed data
serving solutions with innovations targeted at taking advantage of growing flash
speeds. Unfortunately their performance is hampered by the need to maintain and
replicate a write ahead log to guarantee availability in the face of machine and
storage failures. Cyclone is a replicated log plug-in for key value stores that
systematically addresses various sources of this bottleneck. It uses a small
amount of non-volatile memory directly addressable by the CPU - such as in the
form of NVDIMMs or Intel 3DXPoint - to remove block oriented IO devices such as
SSDs from the critical path for appending to the log. This enables it to address
network overheads using an implementation of the RAFT consensus protocol that is
designed around a userspace network stack to relieve the CPU of the burden of
data copies.  Finally, it provides a way to exploit the parallelism available in commodity NICs.  
Cyclone is able to replicate millions of small updates per second using only commodity 10
gigabit ethernet adapters. As a practical application, we use it to improve the
performance (and availability) of RocksDB, a popular persistent key value store
by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

@@

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in services where high availability and
revenue are interconnected. Key value stores therefore usually incorporate
support for a write ahead log that if replicated and kept durable for every
appended update provides the necessary high availability. Unfortunately, the
write ahead log is a performance achilles heel for these systems, eclipsing much
of the work on improving the performance of the LSM component. To illustrate the
impact of the write ahead log, consider Figure~\ref{fig:problem}. The line
marked 'Rocksdb' shows the performance of Rocksdb without the write ahead
log. The performance when persisting the write ahead log without replicating it
is shown as the line marked 'rocksdb/WAL'. The line marked 'rocksdb/3 way rep.'
is for simply replicating the log three ways without persisting it, using
RAFT~\cite{raft} running over TCP/IP. Either persisting every update to the log
or replicating it using TCP/IP causes performance to drop by an order of
magnitude (note the log scale on the x-axis). It is therefore no surprise that
deployments of Rocksdb often turn off the write ahead log~\cite{samza},
depending on upper layers to provide availability by co-ordinating replicas. On
the other side of the spectrum, key value store research prototypes such as
FloDB~\cite{flodb} turn off the write ahead log to be able to showcase benefits
of sophisticated extensions to LSM data structures.

\begin{figure}
\centering \includegraphics[scale=0.6]{results2/problem.pdf}
\caption{Rocksdb write ahead logging impact}
\label{fig:problem}
\end{figure}

These two performance bottlenecks illustrated in Figure~\ref{fig:problem} are
\emph{not independent}, they cannot be addressed in isolation. In particular,
durability semantics mean that the key-value store cannot respond to the client
until the operation has been persisted to the log at a majority of replicas,
since it must be completed even on a crash and subsequent recovery. Therefore,
addressing the network bottleneck is pointless without also addressing the
storage bottleneck. Cyclone shows how a small amount of non-volatile memory can
be used to address both bottlenecks.

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone entirely
closes the performance gap due to both the storage and the network shown in
Figure~\ref{fig:problem}. We show that Cyclone achieves performance transparent
replication in that it can provide both a persistent write ahead log and
replication without compromising on RocksDB's performance.

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. This is due to the fixed cost of a round trip to the SSD through the IO
interface (NVMe in our case), regardless of the volume of data being
synchronously persisted. Small updates (sum of key and value sizes) are
surprisingly common in observed workloads for key value pairs. For example,
Facebook~\cite{fb_workload_analysis} reports that 90\% of the space allocated in
key value stores is for data items under 500 bytes in size.

Cyclone avoids this overhead by making use of a small amount of directly
attached non-volatile memory (NVM), also called persistent memory, on the
server.  This can be in the form of NVDIMMs (DIMMs with an added ultracapacitor
to dump state to a small amount of attached flash) -or- newer and potentially
cheaper (but slower) forms of persistent memory~\cite{pmfs}.
%For current deployments, NVDIMMs are significantly more expensive than standard
%DIMMs, which means that they will %likely be available only in smaller
%quantities compared to flash.
In cases where the log does not entirely fit in NVM, we periodically drain the
NVM log into a log placed on a standard SSD on the system.

The second key contribution of this work is in showing that the throughput
bottlenecks in the network stack can be mitigated entirely in software, once
block IO is removed from the critical path. Existing work on log replication
using state machine replication protocols such as Paxos~\cite{paxos} and
RAFT~\cite{raft} has not - thus far - addressed high performance replication in
the local area network of a datacenter using commodity networks.  Research
focusing on addressing network latencies by reducing the number of
network hops with new protocols such as Fast Paxos~\cite{fast-paxos} or network
switch modifications such as No Paxos~\cite{nopaxos} do not address throughput
bottlenecks due to network stack related inefficiencies in the participating
nodes of the quorum -- a more urgent problem with local area replication. On the
other hand, work done to address network related inefficiencies start with the
assumption that the bottleneck to good throughput on the network is the CPU and
therefore one must either resort to network offload mechanisms such as RDMA~\cite{dare,
  farm, faast} or offload the protocol entirely to an
FPGA~\cite{consensus_box}. We demonstrate in this paper that this is not the
case. Rather, once persistence is provided by directly attached non-volatile
memory, thereby eliminating block IO from the persistence step in replication
protocols, the problem becomes akin to multicasting the same log entry
from the leader to all followers. We design Cyclone's network stack around well
known principles in the software packet switching community. In particular, a
careful implementation of a consensus protocol that relieves the CPU of the
responsibility of data movement permits a high performance implementation
entirely in software that addresses the order of magnitude network related
performance gap in Figure~\ref{fig:problem} using only commodity 10 Gigabit
ethernet.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. The third
key contribution of this work is to show how one can exploit this parallelism in
Cyclone automatically creates and manages multiple logs, mapping them to the 
available parallelism in the system via multiple instances of the same consensus 
protocol.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs in sync
across the replicas. On failure of the leader replica, a new leader is
automatically elected -- RAFT ensures that the new leader has the most up to
date log. The Cyclone client library automatically locates the new leader on a
failover, resulting in only a brief interruption in service for
clients.  Cyclone guarantees that the order of operations to any particular key 
are linearizable. A key sees the exact same sequence of update operations 
(including deletes) on all replicas. These per-key guarantees follow immediately 
from the properties of the RAFT log and by requiring clients to perform a quorum 
read (logged as an entry in the RAFT log).

In addition, Cyclone also provides clients with a weak read that does not require a 
quorum read thereby requiring only one round trip to a server.  A weak read of a key sees a 
prefix of the linearized sequence of updates to a key that includes all successful update operations made 
by the client before it issues the weak read. To do a weak read, 
clients maintain the last known term of the RAFT leader locally and ensure they always talk
to a RAFT leader that has at least the same term. This ensures that clients see their own writes
avoiding split brain problems where a network partition allows clients to
talk to old RAFT leaders after having committed changes to a new RAFT leader. In addition, a weak 
read waits until all pending RAFT log entries have been committed and applied before executing. 
This ensures that in the event of a failover, any previous committed updates from the same client 
have been applied and are visible to the weak read. A weak read is not linearizable as the RAFT 
leader responding to the weak read might actually no longer be the leader - a new one having been 
elected without it realizing, perhaps due to a network partition. This means that while a weak read 
from a client sees its own writes, in some rare cases it might not see later writes by other clients.

@@

\section{Storage}
\label{sec:storage}
Storage devices such as flash-based SSDs export a block IO interface. Appending
a small update to the write-ahead log requires a synchronous access to the
SSD. Even with write coalescing using a write cache on the SSD, the round trip
time to a current generation NVMe SSD such as the one we use in this
paper~\cite{ssd_spec}, is of the order of 20 us, limiting the throughput to
under 50K ops/sec and resulting in a high baseline latency at even low load (as
shown in Figure~\ref{fig:problem}). This problem leads many system designers to
under-provision key-value store shards ensuring that it operates at moderate to
high load, using group-commit to batch additions to the log. They therefore pay
a price in latency to ensure the bandwidth to storage is fully utilized.

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU, rather than being
placed behind a block oriented IO interface. Log appends are done to the ``NVM
log'' placed in this non-volatile memory. However, directly attached persistent 
memory in the form of NVDIMMs today or new memory technologies in the
future~\cite{pmfs} will be of lesser capacity than traditional (NAND based)
flash available through an
SSD. Providing the same capacity as existing flash based logs for key value
stores therefore requires us to provide a second level of the log
placed on a flash SSD (called flashlog).  Entries are drained from
the NVM log to this flashlog in conveniently large units that are a multiple of
4KB -- the optimum IO unit for flash SSDs. This ensures we get the best possible
throughput from a flash device without paying the price of synchronous IO for
small key-value pairs. At the same time, the amount of non-volatile memory
required is small enough to not add undue cost to the server. This is in contrast to
systems like FARM~\cite{farm} that require both the key value store and write ahead log
to be held entirely in non-volatile memory.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

In this section we demonstrated how Cyclone uses a small amount of non-volatile
memory to improve storage bottlenecks when maintaining a durable log on a single
machine. In the next section, we deal with replicating this log efficiently
across multiple machines.

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the log of
operations across machines. Cyclone provides guarantee to the client that any
operation responded to is durable and its results \emph{available} as long as a
majority of replicas continue to function. Cyclone's failure model is
\emph{fail-recovery}, where nodes that can recover from failure simply continue
to participate in the protocol. This requires RAFT to persist every log entry
\emph{before} it sends it out for replication and before a follower replica
responds to the leader. Storage overheads are therefore not independent of
network overheads, and we must deal with both storage and networking overheads
simultaneously to improve the performance of replication.

@@

\begin{figure}
\small
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Event handling}
\label{fig:control_plane}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[scale=0.6]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure*}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures2/nvm_log.pdf}
\caption{NVM log structure}
\label{fig:nvm_log}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

\begin{figure*}
  \centering
  \small
  \bf
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrier.mask != mask of all participants
        if barrier.failed
         send retry to client
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         send retry to client
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

@@

Both the replica count and payload size can have significant impact on Cyclone's
network performance. First, the number of replicas dictates the outgoing message
rate from the leader replica and therefore increasing the replication factor can
decrease Cyclone's performance. Figure~\ref{fig:replicas} shows the impact of
varying replica count. Using only a single replica cuts out a network round trip
and shows the best unloaded latency (10 us) and peak throughput (near 10M
entries/s). Adding replicas decreases the peak throughput down to around 2M
entries/s with 5 replicas. We note that a number of previous pieces of
work~\cite{faast, farm} use three replicas and therefore we focus on three
replicas for the replicated cases we consider below. The second factor that
dictates Cyclone's performance is the size of the log entry being
replicated. Figure~\ref{fig:payload} shows the effect of increasing the payload
size from zero to 512 bytes. Peak throughput drops from 6M entries/s to
approximately 2M entries/s. At this replication rate, the leader replica needs
to transmit data at approximately 30 Gbit/s. Coupled with the cost of network
headers all four 10 GigE ports are now saturated and therefore Cyclone hits the
network line rate bottleneck at this point. It is worthwhile to compare this to
the case with small updates, where running the consensus protocol was a
bottleneck to reaching line rate.

\begin{figure}
  \includegraphics[scale=0.6]{results2/replicas.pdf}
  \caption{Impact of replica count}
  \label{fig:replicas}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/512.pdf}
  \caption{Impact of payload size}
  \label{fig:payload}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/flashlog.pdf}
  \caption{Adding second level log}
  \label{fig:flashlog}
\end{figure}

@@

\begin{figure}
  \includegraphics[scale=0.6]{results2/flashlog_512.pdf}
  \caption{Second level log - 512 bytes payload}
  \label{fig:flashlog_512}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{results2/rocksdb.pdf}
    \caption{Rocksdb - small updates}
    \label{fig:rocksdb}
  \end{figure}

  \begin{figure}
    \includegraphics[scale=0.6]{results2/rocksdb_256.pdf}
    \caption{Rocksdb - large updates}
    \label{fig:rocksdb_256}
  \end{figure}

  \begin{figure}
    \includegraphics[scale=0.6]{results2/rocksdb_multi.pdf}
    \caption{Rocksdb - batched writes}
    \label{fig:rocksdb_multi}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6]{results2/fb.pdf}
\caption{Rocksdb - read heavy}
\label{fig:fb}
\end{figure}
\begin{figure}
  \includegraphics[scale=0.6]{results2/fb2.pdf}
  \caption{Rocksdb - write heavy}
  \label{fig:fb2}
\end{figure}
\begin{figure}
  \includegraphics[scale=0.6]{results2/failover.pdf}
  \caption{Rocksdb - failover}
  \label{fig:timeline}
\end{figure}

>>>>

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix}

@@

\title{Cyclone: High Availability for Persistent Key Value Stores}

\author{{\rm Amitabha Roy} \\ amitabha.roy@gmail.com 
	     \and {\rm Subramanya R. Dulloor} \\ dulloor@gmail.com}

@@

\begin{abstract}
\vspace{0.05in}
Persistent key value stores are an important component of many
distributed data serving solutions with innovations targeted at taking
advantage of growing flash speeds. Unfortunately their performance is
hampered by the need to maintain and replicate a write ahead log to
guarantee availability in the face of machine and storage
failures. Cyclone is a replicated log plug-in for key value stores
that systematically addresses various sources of this bottleneck. It
uses a small amount of non-volatile memory directly addressable by the
CPU - such as in the form of NVDIMMs or Intel 3DXPoint - to remove
block oriented IO devices such as SSDs from the critical path for
appending to the log. This then enables it to address network
overheads using an implementation of the RAFT consensus protocol that
is designed around a userspace network stack to relieve the CPU of the
burden of data copies.  Finally, it provides a way to exploit the
parallelism available in commodity NICs.  Cyclone is able to replicate
millions of small updates per second using only commodity 10 gigabit
ethernet adapters. As a practical application, we use it to improve
the performance (and availability) of RocksDB, a popular persistent
key value store by an order of magnitude when compared to its own
write ahead log without replication.
\end{abstract}

@@

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in services where high availability and
revenue are interconnected. Key value stores therefore usually incorporate
support for a write ahead log that if replicated and kept durable for every
appended update provides the necessary high availability. Unfortunately, the
write ahead log is a performance achilles heel for these systems, eclipsing much
of the work on improving the performance of the LSM component. To illustrate the
impact of the write ahead log, consider Figure~\ref{fig:problem}. The line
marked 'Rocksdb' shows the performance of Rocksdb without the write ahead
log. The performance when persisting the write ahead log without replicating it
is shown as the line marked 'rocksdb/WAL'. The line marked 'rocksdb/3 way rep.'
is for simply replicating the log three ways without persisting it, using
RAFT~\cite{raft} running over TCP/IP. Either persisting every update to the log
or replicating it using TCP/IP causes performance to drop by an order of
magnitude (note the log scale on the x-axis). It is therefore no surprise that
deployments of Rocksdb often turn off the write ahead log~\cite{samza},
depending on upper layers to provide availability by co-ordinating replicas. On
the other side of the spectrum, key value store research prototypes such as
FloDB~\cite{flodb} turn off the write ahead log to be able to showcase benefits
of extensions to LSM data structures.

\begin{figure}
\centering \includegraphics[scale=0.5]{results2/problem.pdf}
\caption{Rocksdb write ahead logging impact}
\label{fig:problem}
\end{figure}

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone entirely
closes the performance gap due to both the storage and the network shown in
Figure~\ref{fig:problem}.

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs. We show that this entirely eliminates the storage bottleneck due to block IO.

The second key contribution of this work is in showing that the
throughput bottlenecks in the network stack can be mitigated entirely
in software, once block IO is removed from the critical path. Existing
work on log replication makes the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must
either resort to network offload mechanisms such as RDMA~\cite{dare,
  farm, faast} or offload the protocol entirely to an
FPGA~\cite{consensus_box}. We demonstrate in this paper that this is
not the case. Rather, once persistence is provided by directly
attached non-volatile memory, thereby eliminating block IO from the
persistence step in replication protocols, the problem becomes akin to
multicasting the same log entry from the leader to all followers. We
design Cyclone's network stack around well known principles in the
software packet switching community. In particular, a careful
implementation of a consensus protocol that relieves the CPU of the
responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of
magnitude network related performance gap in Figure~\ref{fig:problem}
using only commodity 10 Gigabit ethernet.

Both the storage (NVM, SSD) as well as the NIC expose significant
parallelism that is left unused when Cyclone is used to replicate a
single log. The third key contribution of this work is to show how one
can exploit this parallelism for replicating the write ahead log in
key value stores.  Cyclone automatically creates and manages multiple
logs, mapping them to the available parallelism in the system via
multiple instances of the same consensus protocol.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs
in sync across the replicas. On failure of the leader replica, a new
leader is automatically elected -- RAFT ensures that the new leader
has the most up to date log. The Cyclone client library automatically
locates the new leader on a failover, resulting in only a brief
interruption in service for clients.  Cyclone guarantees that the
order of operations to any particular key are linearizable. A key sees
the exact same sequence of update operations (including deletes) on
all replicas. These per-key guarantees follow immediately from the
properties of the RAFT log and by requiring clients reading a key to
perform a quorum read (logged as an entry in the RAFT log).

In addition, Cyclone also provides allows clients to do a weak read
that returns the current value of the key at the last known leader
replica. Weak reads can see a possibly stale value
(i.e. is not linearizable) but is howevere guaranteed to have seen all
of its own previous writes. On the positive side, weak reads avoid an
extra round trip for quorum consensus.

@@

\section{Storage}
\label{sec:storage}
The storage bottleneck for the write ahead log in key value stores
such as Rocksdb is - surprisingly - not the storage media (i.e. SSD)
itself but rather the interface (i.e block IO). This is a particular
pain point for small writes.  For example,
Facebook~\cite{fb_workload_analysis} reports that 90\% of the space
allocated in key value stores is for data items under 500 bytes in
size.  The SSD used in Figure~\ref{fig:problem}, we achieve 50
Kops/sec to the SSD being limited to a 20us latency on the PCIe
interface. This translates to 25 MB/s for 500 byte key-value
pairs. The SSD is capable of achieving over 1 GB/s of sequential write
bandwidth to an append only log. Clearly the SSDs performance is not
the problem, but rather the block IO interface that imposes the round
trip latency \emph{regardless of the size of the update}.

Cyclone solves this problem by making use of novel persistent main
memory technologies. We assume a small amount of non-volatile memory
directly addressable by the CPU, rather than being placed behind a
block oriented IO interface. Log appends are done to the ``NVM log''
placed in this non-volatile memory. However, directly attached
persistent memory in the form of NVDIMMs today or new memory
technologies in the future~\cite{pmfs} will be of lesser capacity than
traditional (NAND based) flash available through an SSD. Providing the
same capacity as existing flash based logs for key value stores
therefore requires us to provide a second level of the log placed on a
flash SSD (called flashlog).  Entries are drained from the NVM log to
this flashlog in conveniently large units that are a multiple of 4KB
-- the optimum IO unit for flash SSDs. This ensures we get the best
possible throughput from a flash device without paying the price of
synchronous IO for small key-value pairs. At the same time, the amount
of non-volatile memory required is small enough to not add undue cost
to the server. This is in contrast to systems like FARM~\cite{farm}
that require both the key value store and write ahead log to be held
entirely in non-volatile memory.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the
log of operations across machines. The key bottleneck in log
replication in Figure~\ref{fig:problem} is the CPU - the achieved
network throughput at load is well under the capacity of the 10 Gbps
link.  The key to improving CPU utilization for log replication is the
insight that log replcation protocols such as RAFT are oblivious to
the actual contents of the log entry being replicated.  This
observation allowed us to greatly improve replication performance by
building a (near) zero copy implementation of RAFT. Cyclone's
implementation of RAFT does only a small \emph{constant} amount of
work per replicated log entry on the CPU leaving the (commodity) NIC
to do the rest of the heavy lifting.

Cyclone provides the guarantee to the client that any
operation responded to is durable and its results \emph{available} as long as a
majority of replicas continue to function. Cyclone's failure model is
\emph{fail-recovery}, where nodes that can recover from failure simply continue
to participate in the protocol. This requires RAFT to persist every log entry
\emph{before} it sends it out for replication and before a follower replica
responds to the leader. Storage overheads are therefore not independent of
network overheads, and we must deal with both storage and networking overheads
simultaneously to improve the performance of replication.

@@

\begin{figure*}
\begin{minipage}{.5\textwidth}
\scriptsize
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Event handling}
\label{fig:control_plane}
\end{minipage}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[scale=0.3]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{minipage}
\end{figure*}

\begin{figure}
\centering
\includegraphics[scale=0.3]{figures2/nvm_log.pdf}
\caption{NVM log structure}
\label{fig:nvm_log}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

\begin{figure}
  \centering
  \scriptsize
  \bf
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrier.mask != mask of all participants
        if barrier.failed
         send retry to client
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         send retry to client
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure}

@@

Both the replica count and payload size can have significant impact on Cyclone's
network performance. First, the number of replicas dictates the outgoing message
rate from the leader replica and therefore increasing the replication factor can
decrease Cyclone's performance. Figure~\ref{fig:replicas} shows the impact of
varying replica count. Using only a single replica cuts out a network round trip
and shows the best unloaded latency (10 us) and peak throughput (near 10M
entries/s). Adding replicas decreases the peak throughput down to around 2M
entries/s with 5 replicas. We note that a number of previous pieces of
work~\cite{faast, farm} use three replicas and therefore we focus on three
replicas for the replicated cases we consider below. The second factor that
dictates Cyclone's performance is the size of the log entry being
replicated. Figure~\ref{fig:payload} shows the effect of increasing the payload
size from zero to 512 bytes. Peak throughput drops from 6M entries/s to
approximately 2M entries/s. At this replication rate, the leader replica needs
to transmit data at approximately 30 Gbit/s. Coupled with the cost of network
headers all four 10 GigE ports are now saturated and therefore Cyclone hits the
network line rate bottleneck at this point. It is worthwhile to compare this to
the case with small updates, where running the consensus protocol was a
bottleneck to reaching line rate.
\begin{figure*}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/replicas.pdf}
  \caption{Impact of replica count}
  \label{fig:replicas}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/512.pdf}
  \caption{Impact of payload size}
  \label{fig:payload}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog.pdf}
  \caption{Adding second level log}
  \label{fig:flashlog}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog_512.pdf}
  \caption{Second level log - 512 bytes payload}
  \label{fig:flashlog_512}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
    \centering
    \includegraphics[scale=0.45]{results2/rocksdb.pdf}
    \caption{Rocksdb - small updates}
    \label{fig:rocksdb}
  \end{minipage}
  \begin{minipage}{.33\textwidth}
    \includegraphics[scale=0.45]{results2/rocksdb_256.pdf}
    \caption{Rocksdb - large updates}
    \label{fig:rocksdb_256}
  \end{minipage}
  \begin{minipage}{.33\textwidth}
    \includegraphics[scale=0.45]{results2/rocksdb_multi.pdf}
    \caption{Rocksdb - batched writes}
    \label{fig:rocksdb_multi}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
\includegraphics[scale=0.45]{results2/fb.pdf}
\caption{Rocksdb - read heavy}
\label{fig:fb}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/fb2.pdf}
  \caption{Rocksdb - write heavy}
  \label{fig:fb2}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/failover.pdf}
  \caption{Rocksdb - failover}
  \label{fig:timeline}
\end{minipage}
\end{figure*}
